Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-30 latex/article.tex
Working on 1 and 2.

@@

\begin{abstract}\sloppy
  Interactive visualization is a very convenient tool to explore
  complex scientific data or to explore different parameter settings
  for a given processing algorithm. In this article, we present a tool
  to efficiently explore the curvature tensor on the boundary of
  potentially large and dynamic digital objects (mean and gaussian curvature,
  principal curvatures, principal directions and normal vector
  field). More precisely, we combine a fully parallel pipeline on GPU
  to extract an adaptive triangulated isosurface of the digital
  object, with a curvature tensor estimation at each surface point
  based on integral invariants. Integral invariants being parametrized
  by a given ball radius, our proposal allows us to explore
  interactively different radii and thus select the appropriate scale
  to which the computation is performed and visualized.

@@

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions, \ldots) estimation at each point of the digital
surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.
Furthermore,
most curvature estimators require a parameter fixing the scale
to which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies noise
level on the object surface. As a consequence, exploring interactively
different values of such parameter is mandatory.
\vspace{0.2cm}

@@

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching Cubes approach by \textsc{Lorensen} and
\textsc{Cline}~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by \textsc{Lengyel} \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.
%However, it
%was only developed on CPU to visualize in real-time static or near
%static data, combining octree with new Marching Cubes configurations
%to generate adaptive meshes.
It combines an octree space partitioning with new Marching Cubes configurations
to generate adaptive meshes on the CPU from
%, but was only developed on the CPU to visualize
static or near static data.
In this paper we propose a full GPU pipeline
%inspired from this method.
inspired from this method,
%using a GPU-friendly Lod criterion
to maintain a view dependent triangulation.
%It leaves the CPU free for other tasks and avoids synchronization with the GPU.
%Our high framerate also allows
%Our high framerate allows this method to run in real-time on fully dynamic data.
Our high framerate allows to visualize fully dynamic data in real-time.
%The triangulation is controlled by a GPU LoD criterion
%that ensures that
%triangles project onto more than one pixel, thus removing geometrical
%aliasing.

Curvature estimation on discrete or digital surface has also been
widely investigated. In \cite{CVIU2014}, authors propose digital
versions of Integral Invariant estimators
\cite{Pottmann2007,Pottmann2009} in order to estimate the complete
curvature tensor (mean/Gaussian curvatures, principal curvatures,
principal curvature directions, normal vector field) on digital
surfaces. Such approaches are based on an integration principle using
a spherical kernel of a given radius. Additionally, authors have
demonstrated that these estimators have multigrid convergence
properties. In this article, we also propose an efficient GPU
implementation of such estimators to visualize%, in real-time,
such curvature maps in real-time and %allowing us
to interactively change the %explore different
value of the spherical kernel radius.

@@

In \cite{DGCI2013}, authors have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ with
$\MCard(Y)$, the number of digital points of the intersection between a ball
$\Ball{R/h}{\vx/h}$ of digital radius $R/h$ and $\DigShape$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
They have proved that estimating both 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property \cite{DGCI2013}. For short, when the digital objet becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
associated one on $\Bd{\Shape}$.
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
Setting $R = kh^\frac{1}{3}$ on convex shapes
with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
related to the maximal curvature of the shape, they have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

@@

Additionally, eigenvectors associated to $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ of the covariance matrix correspond to principal curvature
directions $\PrincDirH{1}{R}$ and $\PrincDirH{2}{R}$. The smallest eigenvector
corresponds to the normal direction $\NormalDirH{R}$ at $\vx$. Convergence
results can be found in \cite{ChapterIICurvature}.%%% Changer avec le chapitre de Jaco à un moment

>>>>

\newcommand{\Jaco}[1]{{\color{orange}#1}}

@@

\begin{abstract}\sloppy
  Interactive visualization is a very convenient tool to explore
  complex scientific data or to explore different parameter settings
  for a given processing algorithm. In this article, we present a tool
  to efficiently explore the curvature tensor on the boundary of
  potentially large and dynamic digital objects (mean and gaussian curvatures,
  principal curvatures, principal directions and normal vector
  field). More precisely, we combine a fully parallel pipeline on GPU
  to extract an adaptive triangulated isosurface of the digital
  object, with a curvature tensor estimation at each surface point
  based on integral invariants. Integral invariants being parametrized
  by a given ball radius, our proposal allows us to explore
  interactively different radii and thus select the appropriate scale
  \Jaco{at} which the computation is performed and visualized.

@@

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.  In this work, we focus on
interactive visualization of volumetric data isosurfaces combined with
a curvature tensor (mean/Gaussian curvature, principal curvatures
directions, \ldots) estimation at each point of the digital
surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.
Furthermore,
most curvature estimators require a parameter fixing the scale
\Jaco{at} which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies \Jaco{a scale for analyzing the object surface, and is naturally related to the amount of perturbation on input data}. As a consequence, exploring interactively
different values of such parameter is mandatory.
\vspace{0.2cm}

@@

\sloppy\noindent\textbf{Related works}\quad Extracting and visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching Cubes approach by \textsc{Lorensen} and
\textsc{Cline}~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
\Jaco{triangulated} mesh resolution \Jaco{according to the object geometry or the
viewpoint}. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by \textsc{Lengyel} \textit{et al.}
\cite{lengyel2010voxel} \Jaco{suits best our needs}.
%However, it
%was only developed on CPU to visualize in real-time static or near
%static data, combining octree with new Marching Cubes configurations
%to generate adaptive meshes.
It combines an octree space partitioning with new Marching Cubes configurations
to generate adaptive meshes on CPU from
%, but was only developed on the CPU to visualize
static or near static data.
We propose here a full GPU pipeline
%inspired from this method.
inspired \Jaco{by} this method,
%using a GPU-friendly Lod criterion
\Jaco{that maintains} a view dependent triangulation.
%It leaves the CPU free for other tasks and avoids synchronization with the GPU.
%Our high framerate also allows
%Our high framerate allows this method to run in real-time on fully dynamic data.
Our high framerate allows \Jaco{us} to inspect dynamic 3D data in real-time.
%The triangulation is controlled by a GPU LoD criterion
%that ensures that
%triangles project onto more than one pixel, thus removing geometrical
%aliasing.

Curvature estimation on discrete or digital surface has also been
widely investigated. In \cite{CVIU2014}, authors propose digital
versions of Integral Invariant estimators
\cite{Pottmann2007,Pottmann2009} in order to estimate the complete
curvature tensor (mean/Gaussian curvatures, principal curvatures,
principal curvature directions, normal vector field) on digital
surfaces. Such approaches are based on an integration principle using
a \Jaco{ball} kernel of a given radius. Additionally, authors have
demonstrated that these estimators have multigrid convergence
properties. In this article, we also propose an efficient GPU
implementation of such estimators to visualize
%, in real-time,
such curvature \Jaco{fields} in real-time and
%allowing us
to interactively change the
%explore different
value of the kernel radius.

@@

In \cite{DGCI2013}, authors have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ \Jaco{(we denote such
quantity $\AreaC(Y,h)$ in dimension 2):}
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
They have proved that estimating both 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property \cite{DGCI2013}. For short, when the digital objet becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
associated one on $\Bd{\Shape}$.
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
Setting $R = kh^\frac{1}{3}$ on convex shapes
with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
related to the maximal curvature of the shape, they have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

@@

Additionally, eigenvectors associated to $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ of the covariance matrix \Jaco{are} principal
curvature direction \Jaco{estimators} $\PrincDirH{1}{R}$ and
$\PrincDirH{2}{R}$. The smallest eigenvector corresponds to the normal
direction $\NormalDirH{R}$ at $\vx$. Convergence results can be found
in \cite{ChapterIICurvature}.
%%% Changer avec le chapitre de Jaco à un moment
